---
title: "Getting started with brms.deeprv"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with brms.deeprv}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 7,
                      fig.height = 4, dpi = 110)
```

`brms.deeprv` lets you drop a fast spatial smooth into a brms-style
Poisson / Gaussian regression by adding a `deepRV(...)` term to the
formula. The smooth is computed via a pre-trained MLP decoder rather
than an exact Cholesky of the GP kernel matrix - same posterior shape,
much faster sampling at moderate-to-large grid sizes.

```{r setup, message = FALSE}
library(brms.deeprv)
library(rstan)
set.seed(1)
```

## 1. Pick a decoder from the catalog

The package ships 32 pre-trained decoders covering the cross product
of `(grid_size, kernel)` for `grid_size in {5, 10, 20, 50, 100, 200,
500, 1000}` and `kernel in {matern_1_2, matern_3_2, matern_5_2, rbf}`.

```{r}
dr <- load_deeprv(domain    = "unit_interval",
                  grid_size = 100,
                  kernel    = "matern_3_2")
print(dr)
```

Each decoder owns its training grid (uniform on `[0, 1]` in v0.1)
and a length-scale prior range. Your `ls_prior` must fit inside
`dr$ls_trained_range`.

## 2. Generate a small synthetic example

We'll fake a Poisson count process whose log-rate is a smooth field on
the decoder's 100-point grid.

```{r simulate}
L  <- as.integer(dr$L)
z  <- rnorm(L)
ls <- 0.15  # ground-truth length-scale, well inside dr's [0.01, 1.0]
W1 <- dr$weights$W1; b1 <- dr$weights$b1
W2 <- dr$weights$W2; b2 <- dr$weights$b2
relu <- function(x) pmax(x, 0)
mu_true <- as.numeric(crossprod(W2, relu(crossprod(W1, c(z, ls)) + b1)) + b2)
y <- rpois(L, exp(mu_true))

df <- data.frame(s       = dr$grid_coords,
                 y       = y,
                 obs_idx = seq_len(L))
head(df)
```

In a real workflow you wouldn't synthesise `mu_true` from the decoder;
you'd have a count vector and a coordinate vector. Two helper rules:

- All observations must sit on `dr$grid_coords`. Use
  `rescale_to_unit_interval()` to map your `[xmin, xmax]` into `[0, 1]`,
  then `which_grid_points()` to get `obs_idx`. `snap_to_grid()`
  handles the rounding case when your `s` lives between grid points.
- `ls_prior` is required and must fit inside `dr$ls_trained_range`.

## 3. Fit

```{r fit, results = "hide"}
options(mc.cores = 2)
fit <- deeprv_brm(
  y ~ deepRV(s,
             decoder  = dr,
             obs_idx  = obs_idx,
             ls_prior = prior_uniform(0.05, 0.5)),
  data    = df,
  family  = poisson(),
  chains  = 2,
  iter    = 1000,
  warmup  = 500,
  seed    = 1
)
```

```{r}
print(fit)
```

`fit` is a `deeprv_fit` wrapping the underlying `stanfit`, the
generated Stan code (under `fit$stancode`), the data list passed to
Stan (`fit$standata`), and the decoder.

## 4. Posterior predict / epred

```{r}
ypred <- posterior_predict(fit)        # response-scale Poisson draws
epred <- posterior_epred(fit)          # expected rate exp(eta)
dim(ypred)                              # (n_draws, n_obs)
```

`posterior_predict()` / `posterior_epred()` are S3 methods on
`rstantools`'s generics, so they dispatch naturally after
`library(brms.deeprv)` or `library(rstantools)`.

## 5. conditional_effects: the smooth itself

The single highest-leverage post-processing tool is
`conditional_effects()` — it returns the per-grid-point posterior
mean and credible interval of the decoder's `mu`, ready to plot.

```{r ce, fig.alt = "Posterior mean of the deepRV smooth with 90% CI ribbon"}
ce <- conditional_effects(fit, probs = c(0.05, 0.95))
plot(ce)
```

`conditional_effects()` honours `by = factor` (one panel per group)
and `unit_square` Kronecker decoders (drawn as a viridis heatmap with
white contour overlay).

## Where to go next

- `vignette("spatial-varying-coefficients")` (forthcoming) — pass a
  numeric `by` to `deepRV()` for a spatially-varying coefficient model.
- `vignette("space-time")` (forthcoming) — `deepRV_st()` for a
  random-walk / AR1 / decoder-time space-time field.
- The smooth quality is bounded by the decoder's training MSE. For
  the L=100 RBF decoder shipped with v0.1, MSE ≈ 0.03 — adequate for
  most applied work. If you need tighter approximation, train your own
  via the recipe in `benchmarks/vae/rstan/train_decoders.py`.
